---
title: "draft code"
author: "John Nobles, Adam Gullion, Jonathan Aguilar"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(magrittr)
library(ggplot2)
library(gapminder)
library(ggthemes)
library(stringr)
``` 

# Variables of interest (hitting):

-   launch angle (`launch_angle_avg`)
-   exit velocity (`exit_velocity_avg`)
-   hard-hit rate (`hard_hit_percent`)
-   barrel batted rate (`barrel_batted_rate`)
-   strikeout percentage (`k_percent`)

```{r}
hitting <- read.csv('Qualified Hitting 2015-2024stats.csv')

# glimpse(hitting)

# hitting %>% select(launch_angle_avg)
```

# a

```{r}
hitting %>% 
  group_by(year) %>% 
  summarize(avg_launch_angle=mean(launch_angle_avg)) %>% 
  ggplot(aes(year, avg_launch_angle)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()
```

# b

```{r}
hitting %>% 
  ggplot(aes(as.factor(year), launch_angle_avg)) +
  geom_boxplot(outlier.color='red') +
  theme_bw()
```

# c

```{r}
hitting %>% 
  ggplot(aes(barrel_batted_rate, exit_velocity_avg)) +
  geom_point() +
  geom_smooth()
```

# d

```{r}
hitting %>% 
  ggplot(aes(as.factor(year), barrel_batted_rate)) +
  geom_boxplot(outlier.color='red') +
  theme_bw()
  
hitting %>% 
  group_by(year) %>% 
  summarize(avg_bbr=mean(barrel_batted_rate)) %>% 
  ggplot(aes(year, avg_bbr)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()
```

# e

```{r}
hitting %>% 
  ggplot(aes(as.factor(year), exit_velocity_avg)) +
  geom_boxplot(outlier.color='red') +
  theme_bw()

hitting %>% 
  group_by(year) %>% 
  summarize(avg_exitv=mean(exit_velocity_avg)) %>% 
  ggplot(aes(year, avg_exitv)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()
```

# f

```{r}
hitting %>% 
  ggplot(aes(as.factor(year), hard_hit_percent)) +
  geom_boxplot(outlier.color='red') +
  theme_bw()

hitting %>% 
  group_by(year) %>% 
  summarize(avg_hhp=mean(hard_hit_percent)) %>% 
  ggplot(aes(year, avg_hhp)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()
```

# g

```{r}
hitting %>% 
  ggplot(aes(as.factor(year), k_percent)) +
  geom_boxplot(outlier.color='red') +
  theme_bw()

hitting %>% 
  group_by(year) %>% 
  summarize(avg_kp=mean(k_percent)) %>% 
  ggplot(aes(year, avg_kp)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()
```

# Variables of interest (pitching)

-   fast ball average velocity (`ff_avg_speed`)
-   fast ball spin (`ff_avg_spin`)
-   fast ball induced vertical break (`ff_avg_break_z_induced`)
-   

```{r}
pitching <- read.csv('QPitchingstats.csv')

glimpse(pitching)
```

# h _ fast-ball spin correlation to k_percent

```{r}

pitching %>% 
  ggplot(aes(ff_avg_spin, k_percent)) +
  geom_point() +
  geom_smooth()

pitching %>% 
  ggplot(aes(ff_avg_speed, k_percent)) +
  geom_point() +
  geom_smooth()
```

# I ff quality vs OPS

```{r}
pitching %>% 
  ggplot(aes(ff_avg_speed, on_base_plus_slg)) +
  geom_point() +
  geom_smooth()

pitching %>% 
  ggplot(aes(ff_avg_spin, on_base_plus_slg)) +
  geom_point() +
  geom_smooth()
```

# J ff vs ex velo& La

```{r}
pitching %>% 
  ggplot(aes(ff_avg_speed, exit_velocity_avg)) +
  geom_point() +
  geom_smooth()

pitching %>% 
  ggplot(aes(ff_avg_spin, exit_velocity_avg)) +
  geom_point() +
  geom_smooth()


pitching %>% 
  ggplot(aes(ff_avg_spin, launch_angle_avg)) +
  geom_point() +
  geom_smooth()

pitching %>% 
  ggplot(aes(ff_avg_speed, launch_angle_avg)) +
  geom_point() +
  geom_smooth()
```

```{r}
pitching %>% 
  ggplot(aes(ff_avg_spin, hard_hit_percent)) +
  geom_point() +
  geom_smooth()


pitching %>% 
  ggplot(aes(ff_avg_speed, hard_hit_percent)) +
  geom_point() +
  geom_smooth()

```

# k ff vs whiff%

```{r}
pitching %>% 
  ggplot(aes(ff_avg_speed, whiff_percent)) +
  geom_point() +
  geom_smooth()


pitching %>% 
  ggplot(aes(ff_avg_spin, whiff_percent)) +
  geom_point() +
  geom_smooth()

```

# L year vs ip

```{r}
pitching %>% 
  ggplot(aes(year, p_formatted_ip)) +
  geom_point() +
  geom_smooth()



pitching %>%
  ggplot(aes(year, p_formatted_ip)) +
  geom_col()

pitching %>%
  group_by(year) %>%
  summarize(avg_p_formatted_ip = mean(p_formatted_ip)) %>%
  ggplot(aes(year, avg_p_formatted_ip)) +
  geom_col()


pitching %>%
  group_by(year) %>%
  summarize(avg_p_formatted_ip = mean(p_formatted_ip, na.rm = TRUE))

summary(pitching$p_formatted_ip)
summary(pitching$year)


pitching %>%
  filter(year != 2020) %>% # Exclude rows where the year is 2020
  group_by(year) %>%
  summarize(avg_p_formatted_ip = mean(p_formatted_ip, na.rm = TRUE))


pitching %>%
  filter(year != 2020) %>% # Remove year 2020
  group_by(year) %>%
  summarize(avg_p_formatted_ip = mean(p_formatted_ip, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_p_formatted_ip)) +
  geom_point() +
  theme_minimal()+ geom_smooth()

```

# M ff vs sweetspot

```{r}
pitching %>% 
  ggplot(aes(ff_avg_spin, sweet_spot_percent)) +
  geom_point() +
  geom_smooth()


pitching %>% 
  ggplot(aes(ff_avg_speed, sweet_spot_percent)) +
  geom_point() +
  geom_smooth()
```

# N (Sticky Stuff)

```{r}
pitching%>% 
  group_by(year) %>% 
  summarize(k_percent=mean(k_percent)) %>% 
  ggplot(aes(year, k_percent)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()


pitching %>% summarize(ff_avg_spin = mean(ff_avg_spin, na.rm = TRUE))
plot_data <- pitching %>%
  group_by(year) %>%
  summarize(ff_avg_spin = mean(ff_avg_spin, na.rm = TRUE))
print(plot_data)


pitching %>%
  gro
#June 21, 2021 - MLB Cracks down on sticky stuff


```up_by(year) %>%
  summarize(ff_avg_spin = mean(ff_avg_spin, na.rm = TRUE)) %>%
  ggplot(aes(year, ff_avg_spin)) +
  geom_point() +
  theme_minimal() +
  geom_smooth()

#Sticky Stuff
