---
title: "John's Section"
author: "John Nobles, Adam Gullion, Jonathan Aguilar"
date: "`r Sys.Date()`"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

# Variables of interest (hitting):

-   launch angle (`launch_angle_avg`)
-   exit velocity (`exit_velocity_avg`)
-   hard-hit rate (`hard_hit_percent`)
-   barrel batted rate (`barrel_batted_rate`)
-   strikeout percentage (`k_percent`)
- and more

```{r include=FALSE}
hitting <- read.csv('Qualified Hitting 2015-2024stats.csv')
```

# Boxplots of strikeout percent by year
- Increase in strikeout percentage from 2015 through 2020 (Spider tack)
- 2 year decline after 2020 (crackdown on Spider tack)
- In 2023 & 2024, median % rose above 20% (result of statcast)

```{r echo=FALSE}
hitting %>% 
  ggplot(aes(as.factor(year), k_percent)) +
  geom_boxplot(outlier.color='red') +
  labs(title='Strikeout Percentage by Year',
       x='Year',
       y='Strikeout %') +
  theme_test()
```

# Average strikeout by Year Trend Line
- Taking a look at the average strikeout percentage by year and a trend line for this data, we can see how our analysis from the previous slide is backed up.

```{r echo=FALSE}
hitting %>% 
  group_by(year) %>% 
  summarize(avg_kp=mean(k_percent)) %>% 
  ggplot(aes(year, avg_kp)) +
  geom_point() +
  theme_test() +
  geom_smooth(method = 'loess', formula = 'y ~ x') +
  scale_x_continuous(breaks=c(2015:2024)) +
  labs(title='Average Strikeout % by Year',
       x='Year',
       y='Strikeout %')
```

```{r echo=FALSE}
pitching <- read.csv('QPitchingstats.csv')
<<<<<<< HEAD
=======

glimpse(pitching)        
>>>>>>> 54b596733d893dc57eee182e54ae9d58eb0c0015
```

# A graph of the boxplots for hard hitting percentage by year
- Layered on top (red dots) is the **average** hard hit percentage for each year.

```{r echo=FALSE}
pitching %>% group_by(year) %>% mutate(avg_hhp_by_year=mean(hard_hit_percent)) %>% 
  ggplot(aes(as.factor(year), hard_hit_percent)) + 
  geom_boxplot() + 
  geom_point(aes(as.factor(year), avg_hhp_by_year), color='red') +
  theme_test() +
  labs(title='Hard hitting percentage by year',
       x='Year',
       y='Hard Hit %',
       color='Average HH% per year',
       legend='a')
```

# Zooming in to *just* the average hard hit percent for each year, we can see a clearer correlation.

```{r echo=FALSE}
pitching %>% group_by(year) %>% mutate(avg_hhp_by_year=mean(hard_hit_percent)) %>% 
  ggplot(aes(year, avg_hhp_by_year)) +
  geom_point() +
  geom_smooth(formula='y ~ x', method='lm') +
  theme_test() +
  labs(title='Hard hitting percentage by year',
       x='Year',
       y='Hard Hit %',
       color='Average HH% per year',
       legend='a') +
  scale_x_continuous(breaks=c(2015:2024))
```

# A violin graph of this data.
- We continue to see the upward trend in hard hit percentage
- What's with the elongated violin in 2020? As sample size decreases, standard deviation increases!

```{r echo=FALSE}
pitching %>% ggplot(aes(as.factor(year), hard_hit_percent,  fill=as.factor(year))) +
  geom_violin() +
  theme_test() +
  labs(title='HH% over Time',
       x='Year',
       y='Hard Hit %')
```

# What is this correlation and how significant is it?
- By conducting a linear regression model, on this data, we see:

```{r echo=FALSE}
summary(lm(hard_hit_percent ~ year, data=hitting))
```

- According to the model, hard hit percentage increased on average .75% per year.
- This increase is associated with a very low p-value, indicating it is significant.
