---
title: "Jonathan's Section"
author: "John Nobles, Adam Gullion, Jonathan Aguilar"
date: "2025-04-15"
output: slidy_presentation
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```
```{r}
hitting <- read.csv('Qualified Hitting 2015-2024stats.csv')
pitching <- read.csv('QPitchingstats.csv')
```

# Hit avarages over the years
- We see slight decrease, but not much change.
- Dip in 2020 due to covid.

```{r}
hits_2015 = hitting %>% 
  filter(year == 2015) %>% select(hit) %>% unlist() %>% mean()

hits_2016 = hitting %>% 
  filter(year == 2016) %>% select(hit) %>% unlist() %>% mean()

hits_2017 = hitting %>% 
  filter(year == 2017) %>% select(hit) %>% unlist() %>% mean()

hits_2018 = hitting %>% 
  filter(year == 2018) %>% select(hit) %>% unlist() %>% mean()

hits_2019 = hitting %>% 
  filter(year == 2019) %>% select(hit) %>% unlist() %>% mean()

hits_2020 = hitting %>% 
  filter(year == 2020) %>% select(hit) %>% unlist() %>% mean()

hits_2021 = hitting %>% 
  filter(year == 2021) %>% select(hit) %>% unlist() %>% mean()

hits_2022 = hitting %>% 
  filter(year == 2022) %>% select(hit) %>% unlist() %>% mean()

hits_2023 = hitting %>% 
  filter(year == 2023) %>% select(hit) %>% unlist() %>% mean()

hits_2024 = hitting %>% 
  filter(year == 2024) %>% select(hit) %>% unlist() %>% mean()

Hits = c(hits_2015,hits_2016,hits_2017,hits_2018,hits_2019,hits_2020,hits_2021,hits_2022,hits_2023,hits_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)

data <- data.frame(years,Hits)

# Plot

suppressWarnings(
  ggplot(data,aes(x=years, y=Hits)) + scale_x_discrete(limits = years) + ylim(0,200) + geom_line()
)
```


# Launch angle average over the years
- Refers to the average vertical angle at which their batted balls leave the bat.

```{r}
abs_2015 = hitting %>% 
  filter(year == 2015) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2016 = hitting %>% 
  filter(year == 2016) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2017 = hitting %>% 
  filter(year == 2017) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2018 = hitting %>% 
  filter(year == 2018) %>% select(launch_angle_avg) %>%  unlist() %>% mean()

abs_2019 = hitting %>% 
  filter(year == 2019) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2020 = hitting %>% 
  filter(year == 2020) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2021 = hitting %>% 
  filter(year == 2021) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2022 = hitting %>% 
  filter(year == 2022) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2023 = hitting %>% 
  filter(year == 2023) %>% select(launch_angle_avg) %>% unlist() %>% mean()

abs_2024 = hitting %>% 
  filter(year == 2024) %>% select(launch_angle_avg) %>% unlist() %>% mean()


launch_angle_avg = c(abs_2015,abs_2016,abs_2017,abs_2018,abs_2019,abs_2020,abs_2021,abs_2022,abs_2023,abs_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)


data <- data.frame(years,launch_angle_avg)

# Plot

suppressWarnings(
  ggplot(data,aes(x=years, y=launch_angle_avg)) + ylab("Launch angle avarge") + scale_x_discrete(limits = years) + ylim(10,14) + geom_line()
)
```

# Flyball percentage over the years
- Fly ball refers to a batted ball that travels on an arcing trajectory into the outfield.

```{r}
abs_2015 = hitting %>% 
  filter(year == 2015) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2016 = hitting %>% 
  filter(year == 2016) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2017 = hitting %>% 
  filter(year == 2017) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2018 = hitting %>% 
  filter(year == 2018) %>% select(flyballs_percent) %>%  unlist() %>% mean()

abs_2019 = hitting %>% 
  filter(year == 2019) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2020 = hitting %>% 
  filter(year == 2020) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2021 = hitting %>% 
  filter(year == 2021) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2022 = hitting %>% 
  filter(year == 2022) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2023 = hitting %>% 
  filter(year == 2023) %>% select(flyballs_percent) %>% unlist() %>% mean()

abs_2024 = hitting %>% 
  filter(year == 2024) %>% select(flyballs_percent) %>% unlist() %>% mean()


flyballs_percent = c(abs_2015,abs_2016,abs_2017,abs_2018,abs_2019,abs_2020,abs_2021,abs_2022,abs_2023,abs_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)


data <- data.frame(years,flyballs_percent)

# Plot

suppressWarnings(
  ggplot(data,aes(x=years, y=flyballs_percent)) + ylab("Flyball percentage ") + scale_x_discrete(limits = years) + geom_line()

)
```

# Line drives percentage over the years
- Line drives are batted balls that travel in a sharp, straight line with minimal arc, usually not far above the ground.

```{r}
hits_2015 = hitting %>% 
  filter(year == 2015) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2016 = hitting %>% 
  filter(year == 2016) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2017 = hitting %>% 
  filter(year == 2017) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2018 = hitting %>% 
  filter(year == 2018) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2019 = hitting %>% 
  filter(year == 2019) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2020 = hitting %>% 
  filter(year == 2020) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2021 = hitting %>% 
  filter(year == 2021) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2022 = hitting %>% 
  filter(year == 2022) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2023 = hitting %>% 
  filter(year == 2023) %>% select(linedrives_percent) %>% unlist() %>% mean()

hits_2024 = hitting %>% 
  filter(year == 2024) %>% select(linedrives_percent) %>% unlist() %>% mean()

linedrives_percent = c(hits_2015,hits_2016,hits_2017,hits_2018,hits_2019,hits_2020,hits_2021,hits_2022,hits_2023,hits_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)

data <- data.frame(years,linedrives_percent)

# Plot
suppressWarnings(
  ggplot(data,aes(x=years, y=linedrives_percent)) + ylab("Line drives %") + scale_x_discrete(limits = years) + ylim(0,50) + geom_line()
)

```



# How is swing percentage effected over the years?
- Swing percentage - The percentage of pitches the batter swings at, regardless if they hit or missed the ball.

```{r}
hits_2015 = hitting %>% 
  filter(year == 2015) %>% select(swing_percent) %>% unlist() %>% mean()



hits_2016 = hitting %>% 
  filter(year == 2016) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2017 = hitting %>% 
  filter(year == 2017) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2018 = hitting %>% 
  filter(year == 2018) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2019 = hitting %>% 
  filter(year == 2019) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2020 = hitting %>% 
  filter(year == 2020) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2021 = hitting %>% 
  filter(year == 2021) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2022 = hitting %>% 
  filter(year == 2022) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2023 = hitting %>% 
  filter(year == 2023) %>% select(swing_percent) %>% unlist() %>% mean()

hits_2024 = hitting %>% 
  filter(year == 2024) %>% select(swing_percent) %>% unlist() %>% mean()

swing_percent = c(hits_2015,hits_2016,hits_2017,hits_2018,hits_2019,hits_2020,hits_2021,hits_2022,hits_2023,hits_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)

data <- data.frame(years,swing_percent)

# Plot
suppressWarnings(ggplot(data,aes(x=years, y=swing_percent)) + scale_x_discrete(limits = years) + ylab("Swing %") + ylim(0,100) + geom_line()
)
```


# Four seam fast-ball speed vs. avarge spin correlation? 

```{r, warning = FALSE, message = FALSE}
  pitching %>% ggplot(aes(x = ff_avg_speed,y = ff_avg_spin)) + geom_point() + geom_smooth()

```

#

```{r, warning = FALSE, message = FALSE}


pitching %>% ggplot(aes(x = sl_avg_speed, y = sl_avg_spin)) + geom_point() + geom_smooth()


```

#

```{r, warning = FALSE, message = FALSE}
pitching %>% ggplot(aes(x = ch_avg_speed, y = ch_avg_spin)) + geom_point() + geom_smooth()

```

# Perfect (P) games avarage over the years
- A perfect game occurs when a pitcher, or a series of pitchers, retires all 27 opposing batters without allowing any of them to reach base.

```{r}
abs_2015 = pitching %>% 
  filter(year == 2015) %>% select(p_game) %>% unlist() %>% mean()

abs_2016 = pitching %>% 
  filter(year == 2016) %>% select(p_game) %>% unlist() %>% mean()

abs_2017 = pitching %>% 
  filter(year == 2017) %>% select(p_game) %>% unlist() %>% mean()

abs_2018 = pitching %>% 
  filter(year == 2018) %>% select(p_game) %>% unlist() %>% mean()

abs_2019 = pitching %>% 
  filter(year == 2019) %>% select(p_game) %>% unlist() %>% mean()

abs_2020 = pitching %>% 
  filter(year == 2020) %>% select(p_game) %>% unlist() %>% mean()

abs_2021 = pitching %>% 
  filter(year == 2021) %>% select(p_game) %>% unlist() %>% mean()

abs_2022 = pitching %>% 
  filter(year == 2022) %>% select(p_game) %>% unlist() %>% mean()

abs_2023 = pitching %>% 
  filter(year == 2023) %>% select(p_game) %>% unlist() %>% mean()

abs_2024 = pitching %>% 
  filter(year == 2024) %>% select(p_game) %>% unlist() %>% mean()

p_game = c(abs_2015,abs_2016,abs_2017,abs_2018,abs_2019,abs_2020,abs_2021,abs_2022,abs_2023,abs_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)

data <- data.frame(years,p_game)

# Plot
suppressWarnings(
  ggplot(data,aes(x=years, y=p_game)) + scale_x_discrete(limits = years) + ylab("P game") + ylim(0,35) + geom_line()
)
```

# Four seam fast-ball (FF) spin rate avarage over the years
- (FF) is a common type of pitch thrown by pitchers, characterized by its straight, forward movement.

```{r}
abs_2015 = pitching %>% 
  filter(year == 2015) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2016 = pitching %>% 
  filter(year == 2016) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2017 = pitching %>% 
  filter(year == 2017) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2018 = pitching %>% 
  filter(year == 2018) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2019 = pitching %>% 
  filter(year == 2019) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2020 = pitching %>% 
  filter(year == 2020) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2021 = pitching %>% 
  filter(year == 2021) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2022 = pitching %>% 
  filter(year == 2022) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2023 = pitching %>% 
  filter(year == 2023) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

abs_2024 = pitching %>% 
  filter(year == 2024) %>% select(ff_avg_spin) %>% filter(!is.na(ff_avg_spin)) %>% unlist() %>% mean()

ffavgspin = c(abs_2015,abs_2016,abs_2017,abs_2018,abs_2019,abs_2020,abs_2021,abs_2022,abs_2023,abs_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)


data <- data.frame(years,ffavgspin)

# Plot
suppressWarnings(ggplot(data,aes(x=years, y=ffavgspin)) + scale_x_discrete(limits = years) + ylab("(FF) spin rate avarage") + ylim(0,3000) + geom_line()
                 )

```


# Change up (CH) avaerage spin rate over the years
- The "CH" stands for change up, a type of pitch characterized by its low speed and tendency to move away from the hitter.

```{r}
abs_2015 = pitching %>% 
  filter(year == 2015) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2016 = pitching %>% 
  filter(year == 2016) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2017 = pitching %>% 
  filter(year == 2017) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2018 = pitching %>% 
  filter(year == 2018) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2019 = pitching %>% 
  filter(year == 2019) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2020 = pitching %>% 
  filter(year == 2020) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2021 = pitching %>% 
  filter(year == 2021) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2022 = pitching %>% 
  filter(year == 2022) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2023 = pitching %>% 
  filter(year == 2023) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

abs_2024 = pitching %>% 
  filter(year == 2024) %>% select(ch_avg_spin) %>% filter(!is.na(ch_avg_spin)) %>% unlist() %>% mean()

ch_avg_spin = c(abs_2015,abs_2016,abs_2017,abs_2018,abs_2019,abs_2020,abs_2021,abs_2022,abs_2023,abs_2024)
years = c(2015,2016,2017,2018,2019,2020,2021,2022,2023,2024)


data <- data.frame(years,ch_avg_spin)

# Plot
suppressWarnings(
  ggplot(data,aes(x=years, y=ch_avg_spin)) + scale_x_discrete(limits = years) + ylab("(CH) avaerage spin rate") + ylim(0,3000) + geom_line()
)

```
